\bf Visual Inference for Significance and Goodness-of-Fit Testing for a Social Network Model

### 2017-10-30 ASApaper/paper-child.Rnw
respond to hh comments; add to the power discussion

\section{Introduction}
% \st{
% \begin{enumerate}
% \item significance testing: what is it and why is it important
% \item goodness-of-fit testing: what is it and why is it important
% \item difficulty of these two for sna, why do vis inf
% \end{enumerate}
% }
Two of the most important pieces of statistical modeling are significance testing of model parameters and goodness-of-fit tests. In the former, the data are usually assumed to come from a simple model under the null hypothesis, and additional parameters are tested whether they significantly contribute to explaining variability in the data. In the latter, the model of interest is examined to determine how well it fits the data. Both of these aspects of statistical modeling increase greatly in difficulty as the complexity of the model of interest increases. 
%The more complicated the model, the harder it is to determine fit or to include or exclude a parameter.

@@

We propose new methods for significance and goodness-of-fit testing for a set of social network models, stochastic actor-oriented models for dynamic network data \citep{saompaper}. Specifically, we are using \textit{visual inference} in place of traditional statistical methods for social network models, such as Wald tests for significance of parameters and in- and outdegree distribution metrics for determining goodness-of-fit. Visual inference, introduced by \citet{Bujaetal}, allows us to look at the entire dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network such as outdegree, or a $p$-value for a single parameter in the model. In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.

@@

To simulate lineups from the models, we set the parameters to the values given in Table~\ref{tab:fittedvalues} for all parameters within the respective models, with the exception of $\beta_5$. For $\beta_5$, twice the estimated value was used. More detail on why we use twice the fitted value is provided in Section~\ref{sec:setupvd}.\footnote{If you would like to explore the kinds of lineups we use in further detail, please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/}}. To get the simulations, we used the \texttt{siena07} function in \texttt{RSiena} \citep{RSiena}.

\begin{figure}
\includegraphics[width=.9\textwidth]{img/shinyappscreen.png}
\caption{\label{fig:shiny} A screen shot of the web application we created to design our lineup experiment. More details about this application are given in Section~\ref{sec:setupvd}. In the lineup, M4 is the alternative model with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:fittedvalues}. One plot simulated from this model is placed at random among five observations simulated from the null model, M1. Participants of the study are asked to identify the most different plot.  %The other two tabs in the right side of the screen tell you which plot comes from the alternative model, and allow you to view and download the data that is plotted in the lineup.
}
\end{figure}

@@

\subsection{Visual Detection}\label{sec:setupvd}

Visual detection does not fit into the typical statistical testing framework as easily as the significance and goodness-of-fit tests we perform do. Through visual inference, we want to determine at which point an effect becomes noticeable in a SAOM. By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model to a degree that viewers are able to reliably pick out a node-link diagram rendered from data simulated from a model with the effect from a lineup of plots without the effect. We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.

@@

We recruited 250 participants for our experiment through Amazon Mechanical Turk. Each participant was peresented with some brief training material before beginning the experiment. After agreeing to participate, the participants were shown two trial plots, one where the data plot was the most different from the others due to its complex structure, while the other trial included a data plot that was most different from the others due to its very simple structure. Only when participants were able to correctly identify the data plot from the trial lineups, they were allowed to begin the experiment. Each participant was randomly assigned 13 lineups to look at. The were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice. They could select from ``Most simple overall structure," ``Most complex overall structure," or ``Other" and provide their own text description of their reasoning. 
\hh{what about the first 12 plots? 6 parameters, 3 levels, 2 directions}
One of the 13 lineups the participants saw was the true data from the 112th senate shown in Section~\ref{sec:senatedata}. Each participant only saw the data one time in order to avoid bias. Upon completion of the 13 lineups, each participant was paid \$1.75.

@@

All of the parameters we test for significance using the lineup protocol, $\beta_3$ and $\beta_4$, were determined to be statistically significant using Equation~\ref{eq:wald1}. The results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}. \hh{XXX how do we interpret these  visual results? Include the six lineups in an appendix and refer to them from here.}

@@

We see that the $p$-values from visual inference, which are calculated using the \texttt{vinference} \texttt{R} package by \citet{vinference}), are variable. This variability is introduced through the null plots generated from M1, since not all simulations look alike. In addition, the necessarily small number of null plots do not give the viewer as complete of a view of the null model as the usual 19 null plots would. The results of the significance tests for $\beta_3$ and $\beta_4$ are not clear cut. Thus, unlike the Wald-type tests described at the beginning of this section, there is no way to flat out reject or to fail to reject the null hypothesis that the parameter value is 0. We include all of the lineups shown to our participants in the appendix.

@@

The $p$-values were calculated using the \texttt{vinference} package by \citet{vinference}. This package contains methods to calculate \textit{Visual distributions} for lineup experiment data.  The distribution depends on the number of evaluations of a plot, $K$, the size of the lineup, $m$, and the lineup scenario, which here is that each lineup containing the same data and the same set of null plots is shown to $K$ independent observers.  Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$. The lineup that resulted in a failure to reject the null hypothesis is shown in Figure~\ref{fig:failgof}. The null model in this lineup is M4, and the senate data is shown in panel number $3^2 - 7$. However, the panel most participants chose was number four, and the most common reasoning for that choice was that it had the most simple structure.

<<failgof, fig.height=3, fig.cap="The goodness-of-fit lineup that failed to reject the null hypothesis. The null model for this lineup is M4. Only 7 of 20 viewers of this lineup selected the data plot as the most different from the others.">>=
load("data/se112adjmat.RDS")
wave2 <- data.frame(se112adj)
senators <- data.frame(id = colnames(wave2), number = 1:155)
wave2$from <- colnames(wave2)
wave2 <- wave2 %>% gather(to, val, Alan.Stuart.Franken:William.Cowan) %>% filter(val > 0)
wave2 <- merge(wave2, senators, by.x = "from", by.y = "id", all = T)
wave2$from <- senators$number[match(wave2$from, senators$id)]
wave2$to <- senators$number[match(wave2$to, senators$id)]
wave2$from <- paste0("V", wave2$from)
wave2$to <- ifelse(is.na(wave2$to), NA, paste0("V", wave2$to))
wave2 <- wave2 %>% mutate(sim = 1001, model = "data", wave = 1) %>% select(-c(val, number))
dat <- read_csv(paste0("data/jtts_gof_9_2.csv"))
datplot <- unique(dat$ord[dat$sim == 1001])
dat <- dat %>% filter(sim != 1001) %>% 
      bind_rows(wave2) %>% 
      mutate(ord = ifelse(is.na(ord), datplot, ord))
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
@

@@

\st{I reformulated the discussion per our convo today.  }
\hh{XXX we need to go throught the logic of the argument below.}

<<mostsiggof, fig.height = 3, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis. Surprisingly, this another repetition for M4 as the null model.">>=
dat <- read_csv(paste0("data/jtts_gof_9_3.csv"))
datplot <- unique(dat$ord[dat$sim == 1001])
dat <- dat %>% filter(sim != 1001) %>% 
      bind_rows(wave2) %>% 
      mutate(ord = ifelse(is.na(ord), datplot, ord))
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
@

@@

\subsection{Visual Detection}

A summary of the results from our experiment is shown in Figure~\ref{fig:predictglmm}. On the $x$ axis, we plot the value of the parameter of interest, and on the $y$ axis, the proportion of times the plot of data simulated from the alternative model was picked out for each lineup. The results are split into groups based on the sign of the parameter. We can see clear patterns in all of the groups: as the parameter value approaches 0, fewer participants identified the alternative plot.

@@

\hh{this is the power of the visual test XXX}
We futher explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineuptype with a generalized linear mixed model. The response variable, $Y_{i\ell m}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $\ell$. The covariate $x_{\ell 1}$ takes on the value -1 or 1 according to the lineup type code in Table~\ref{tab:experdetail}. In most cases, -1 indicates the parameter value is less than original estimate, while 1 indicates the parameter value is at or above the original estimate. The continuous covariate $x_{\ell 2}$ is the centered and scaled size of the effect of interest from which the alternative data were simulated, the values of which are labeled ``easy", ``medium", and ``hard" in Table~\ref{tab:experdetail} according to how difficult we thought the Turk participants would find each lineup. In Equation~\ref{eq:glmm}, $i \in \{1, 2, 3, 4, 5\}$ corresponding to the effects $\beta_3, \beta_4, \beta_2, \beta_6, \beta_5$, respectively. We include random effects in the model for each lineup, $\delta_{\ell}$ and each participant, $\epsilon_m$, and fit a hierarchical model as follows:
\begin{align}
\begin{split}
Y_{i\ell m} &\sim \text{Bernoulli}(\pi_{i\ell m}) \\
\text{logit}(\pi_{i\ell m}) & = \mu + \alpha_{i} + \theta\mathbb{I}(x_{\ell 1} = 1) + \gamma x_{\ell 2} + \\ & (\alpha\theta)_i \mathbb{I}(x_{\ell 1} = 1) + (\alpha\gamma)_i x_{\ell 2} + (\theta\gamma)\mathbb{I}(x_{\ell 1} = 1)x_{\ell 2} + \\ & (\alpha\theta\gamma)_i\mathbb{I}(x_{\ell 1} = 1) x_{\ell 2} + \delta_{\ell} + \epsilon_{m} \label{eq:glmm} \\
\delta_{\ell} & \sim N(0, \sigma^2_{\delta}) \\
\epsilon_{m} & \sim N(0, \sigma^2_{\epsilon})
\end{split}
\end{align}
The results of fitting this model using \texttt{glmer} from the \texttt{lme4} package are summarized in Table~\ref{tab:glmmests} \citep{lme4}. The baseline condition for this model against which all other experimental conditions are compared is the first condition in Table~\ref{tab:experdetail}, for parameter $\beta_1$ and lineup type -1. The expected value of the link function for this scenario with $\beta_1 = 0$ is $\mu$, and the expected probability that a new observer will pick out the data plot in a new lineup generated from this scenario is $\frac{\exp(\mu)}{1+\exp{\mu}}$. Similarly, for a new lineup generated from the tenth condition in in Table~\ref{tab:experdetail}, with $\beta_6 = 1$, and lineup type 1, the expected value of the link function is $\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4$. The corresponding expected probability that a new observer viewing a new lineup from this scenario is: 
\begin{equation}
\frac{\exp\{\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4\}}{1 + \exp\{\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4\}}
\end{equation}

@@

<<predictglmm, fig.height=6, fig.cap='Predictions from our generalized linear mixed effects model given in Equation~\ref{eq:glmm}. The lines show the expected probability of detecting the alternative data in a lineup of size 6 for new observers of new lineups is plotted on the $y$-axis, and the size of the parameter of interest is on the $x$-axis. The proportions detected by our Turk participants for each lineup group are shown by the points, with the probability of picking out the data plot at random shown by a horizontal line at 1/6. The lineup marked as ``outlier" was removed from modeling. The panel for the reciprocity parameter, $\\beta_2$ is also presented in Figure~\\ref{fig:recipzoom} in more detail.'>>=
newdata <- read_csv("data/newdata_pred_glmm.csv")
newdata$param <- newdata$test_param
newdata$param <- as.factor(newdata$test_param)
levels(newdata$param) <- c("beta[1]", "beta[3]", "beta[4]", "beta[2]", "beta[6]", "beta[5]")
newdata$param <-as.character(newdata$param)
plotdata <- turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
                       type2 = ifelse(type == "one", type2, -1)) %>% 
  group_by(pic_id, size, test_param, sign, type2,alt_model) %>% summarize(
    datapick = mean(datapick)
)
plotdata$param <- plotdata$test_param
plotdata$param <- as.factor(plotdata$param)
levels(plotdata$param) <- c("beta[1]", "beta[3]", "beta[4]", "beta[2]", "beta[6]", "beta[5]")
plotdata$param <- as.character(plotdata$param)
labdat <- data_frame(test_param = "jttp", param = "beta[3]", type2 = -1, x = -5, y = .9, label = "outlier")

@@

<<recipzoom, fig.height=2.5, fig.cap='The top middle panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail. For this parameter, as its value approaches zero, the probability of identifying the alternate data model decreases, then increases, which is noticeably different from the pattern exhibited by the others. Again, a horizontal line is drawn at 1/6, the chance of selecting the data plot at random.'>>=
ggplot() + 
  geom_line(data = newdata %>% filter(test_param == "recip"), aes(x = size, y = predictfinal, color = as.factor(type2))) + 
  geom_point(data = plotdata %>% filter(test_param == "recip"), aes(x = size, y = datapick, color = as.factor(type2))) + 
  scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
  geom_hline(yintercept = 1/6, linetype='dashed') + 
  facet_wrap(~type2, scales = 'free_x') + 
  labs(x = "Parameter value (recip)", y = "Expected probability of detection\nin a lineup of size 6") + 
  ThemeNoNet + 
  theme(legend.position = 'bottom')
@

@@

<<beta41, fig.height = 3, fig.cap='In our experiment, 52.8\\% of viewers of this plot selected the plot from the alternative model, M4. The ``reverse" of this lineup is given in Figure~\\ref{fig:beta4neg1}, where 41.4\\% of viewers selected the plot from the alternative model, M1. Here, the alternative plot is $\\sqrt{25} - 3$.'>>=
dat <- read_csv("data/jtts_pos_easy_3.csv")
# "answer" is 2
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
@

<<beta4neg1,fig.height=3, fig.cap='In our experiment, 41.4\\% of viewers of this plot selected the plot from the alternative model, M1. The ``reverse" of this lineup is given in Figure~\\ref{fig:beta41}, where 52.8\\% of viewers selected the plot from the alternative model, M1. Here, the alternative plot is $\\sqrt{25} - 1$.' >>=
dat <- read_csv("data/jtts_neg_med_2.csv")
# 'answer' is 4
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
@

@@

>>>>

\section{Introduction}
% \st{
% \begin{enumerate}
% \item significance testing: what is it and why is it important
% \item goodness-of-fit testing: what is it and why is it important
% \item difficulty of these two for sna, why do vis inf
% \end{enumerate}
% }
\st{Three} of the most important pieces of statistical modeling are significance testing of model parameters, goodness-of-fit tests, and power of a test. In the first, the data are usually assumed to come from a simple model under the null hypothesis, and additional parameters are tested whether they significantly contribute to explaining variability in the data. In the second, the model of interest is examined to determine how well it fits the data. \st{In the final, the ability of the  hypothesis test in question to detect the difference between the null and alternative hypothesis is determined.}  All three of these aspects of statistical modeling increase greatly in difficulty as the complexity of the model of interest increases. 
%The more complicated the model, the harder it is to determine fit or to include or exclude a parameter.

@@

We propose new methods for significance and goodness-of-fit testing and power calculation of these for a set of social network models, stochastic actor-oriented models for dynamic network data \citep{saompaper}. Specifically, we are using \textit{visual inference} in place of traditional statistical methods for social network models, such as Wald tests for significance of parameters and in- and outdegree distribution metrics for determining goodness-of-fit. Visual inference, introduced by \citet{Bujaetal}, allows us to look at the entire dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network such as outdegree, or a $p$-value for a single parameter in the model. %In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.

@@

To simulate lineups from the models, we set the parameters to the values given in Table~\ref{tab:fittedvalues} for all parameters within the respective models, with the exception of $\beta_5$. For $\beta_5$, twice the estimated value was used. More detail on why we use twice the fitted value is provided in Section~\ref{sec:setupvp}.\footnote{If you would like to explore the kinds of lineups we use in further detail, please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/}}. To get the simulations, we used the \texttt{siena07} function in \texttt{RSiena} \citep{RSiena}.

\begin{figure}
\includegraphics[width=.9\textwidth]{img/shinyappscreen.png}
\caption{\label{fig:shiny} A screen shot of the web application we created to design our lineup experiment. More details about this application are given in Section~\ref{sec:setupvp}. In the lineup, M4 is the alternative model with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:fittedvalues}. One plot simulated from this model is placed at random among five observations simulated from the null model, M1. Participants of the study are asked to identify the most different plot.  %The other two tabs in the right side of the screen tell you which plot comes from the alternative model, and allow you to view and download the data that is plotted in the lineup.
}
\end{figure}

@@

\subsection{Visual Power}\label{sec:setupvp}

%Visual detection does not fit into the typical statistical testing framework as easily as the significance and goodness-of-fit tests we perform do. 
Through visual inference, we want to determine at which point an effect becomes noticeable in a SAOM. By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model to a degree that viewers are able to reliably pick out a node-link diagram rendered from data simulated from a model with the effect from a lineup of plots without the effect. \st{This is a way to determine the power of the visual test.}  We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.

@@

We recruited 250 participants for our experiment through Amazon Mechanical Turk. Each participant was peresented with some brief training material before beginning the experiment. After agreeing to participate, the participants were shown two trial plots, one where the data plot was the most different from the others due to its complex structure, while the other trial included a data plot that was most different from the others due to its very simple structure. Only when participants were able to correctly identify the data plot from the trial lineups, they were allowed to begin the experiment. Each participant was randomly assigned 13 lineups to look at. The were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice. They could select from ``Most simple overall structure," ``Most complex overall structure," or ``Other" and provide their own text description of their reasoning. 
\st{Twelve of the 13 lineups that the participants saw were used for the significance testing and the visual power methods discussed in Sections~\ref{sec:setupst} and \ref{sec:setupvp}. The six parameters, $\beta_1, \dots, \beta_6$ were set to three different values according to how difficult we thought picking the data plot from the lineup would be for our participants, and were also set to be less than or greater than the initial estimate, creating the lineup type variable.}\hh{what about the first 12 plots? 6 parameters, 3 levels, 2 directions}
One of the 13 lineups the participants saw was the true data from the 112th senate shown in Section~\ref{sec:senatedata}. Each participant only saw the data one time in order to avoid bias. Upon completion of the 13 lineups, each participant was paid \$1.75.

@@

All of the parameters we test for significance using the lineup protocol, $\beta_3$ and $\beta_4$, were determined to be statistically significant using Equation~\ref{eq:wald1}. The results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}. \hh{XXX how do we interpret these  visual results? Include the six lineups in an appendix and refer to them from here.} \st{XXX should be taken care of in next paragraph now.}

@@

We see that the $p$-values from visual inference, which are calculated using the \texttt{vinference} \texttt{R} package by \citet{vinference}), are highly variable. This variability is introduced through the null plots generated from M1, since not all simulations look alike. In addition, the necessarily small number of null plots do not give the viewer as complete of a view of the null model as the usual 19 null plots would. The results of the significance tests for $\beta_3$ and $\beta_4$ are not clear cut. Thus, unlike the Wald-type tests described at the beginning of this section, there is no way to flat out reject or to fail to reject the null hypothesis that the parameter value is 0. We include all of the lineups shown to our participants in the appendix.

@@

The $p$-values were calculated using the \texttt{vinference} package by \citet{vinference}. This package contains methods to calculate \textit{Visual distributions} for lineup experiment data.  The distribution depends on the number of evaluations of a plot, $K$, the size of the lineup, $m$, and the lineup scenario, which here is that each lineup containing the same data and the same set of null plots is shown to $K$ independent observers. \st{The visual inference family of distributions is similar to the binomial distribution, but takes the dependency among the $m$ plots in a signle lineup shown to multiple viewers into account. } Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$. The lineup that resulted in a failure to reject the null hypothesis is shown in Figure~\ref{fig:failgof}. The null model in this lineup is M4, and the senate data is shown in panel number $3^2 - 7$. However, the panel most participants chose was number four, and the most common reasoning for that choice was that it had the most simple structure.

<<failgof, fig.height=4, fig.cap="The goodness-of-fit lineup that failed to reject the null hypothesis. The null model for this lineup is M4. Only 7 of 20 viewers of this lineup selected the data plot as the most different from the others.">>=
load("data/se112adjmat.RDS")
wave2 <- data.frame(se112adj)
senators <- data.frame(id = colnames(wave2), number = 1:155)
wave2$from <- colnames(wave2)
wave2 <- wave2 %>% gather(to, val, Alan.Stuart.Franken:William.Cowan) %>% filter(val > 0)
wave2 <- merge(wave2, senators, by.x = "from", by.y = "id", all = T)
wave2$from <- senators$number[match(wave2$from, senators$id)]
wave2$to <- senators$number[match(wave2$to, senators$id)]
wave2$from <- paste0("V", wave2$from)
wave2$to <- ifelse(is.na(wave2$to), NA, paste0("V", wave2$to))
wave2 <- wave2 %>% mutate(sim = 1001, model = "data", wave = 1) %>% select(-c(val, number))
dat <- read_csv(paste0("data/jtts_gof_9_2.csv"))
datplot <- unique(dat$ord[dat$sim == 1001])
dat <- dat %>% filter(sim != 1001) %>% 
      bind_rows(wave2) %>% 
      mutate(ord = ifelse(is.na(ord), datplot, ord))
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
@

@@

\hh{XXX we need to go throught the logic of the argument below.}\st{ XXX I reformulated the discussion per our convo today.  }

<<mostsiggof, fig.height = 4, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis. Surprisingly, this another repetition for M4 as the null model.">>=
dat <- read_csv(paste0("data/jtts_gof_9_3.csv"))
datplot <- unique(dat$ord[dat$sim == 1001])
dat <- dat %>% filter(sim != 1001) %>% 
      bind_rows(wave2) %>% 
      mutate(ord = ifelse(is.na(ord), datplot, ord))
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
@

@@

\subsection{Visual Power}

A summary of the results from our experiment is shown in Figure~\ref{fig:predictglmm}. On the $x$ axis, we plot the value of the parameter of interest, and on the $y$ axis, the proportion of times the plot of data simulated from the alternative model was picked out for each lineup. The results are split into groups based on the value of the parameter and the lineup type. We can see clear patterns in nearly all of the groups: as the parameter value approaches 0, fewer participants identified the alternative plot.

@@

\hh{this is the power of the visual test XXX}
We futher explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineuptype with a generalized linear mixed model \st{that provides us with an estimate of the power of the visual test.} The response variable, $Y_{i\ell m}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $\ell$. The covariate $x_{\ell 1}$ takes on the value -1 or 1 according to the lineup type code in Table~\ref{tab:experdetail}. In most cases, -1 indicates the parameter value is less than original estimate, while 1 indicates the parameter value is at or above the original estimate. The continuous covariate $x_{\ell 2}$ is the centered and scaled size of the effect of interest from which the alternative data were simulated, the values of which are labeled ``easy", ``medium", and ``hard" in Table~\ref{tab:experdetail} according to how difficult we thought the Turk participants would find each lineup. In Equation~\ref{eq:glmm}, $i \in \{1, 2, 3, 4, 5\}$ corresponding to the effects $\beta_3, \beta_4, \beta_2, \beta_6, \beta_5$, respectively. We include random effects in the model for each lineup, $\delta_{\ell}$ and each participant, $\epsilon_m$, and fit a hierarchical model as follows:
\begin{align}
\begin{split}
Y_{i\ell m} &\sim \text{Bernoulli}(\pi_{i\ell m}) \\
\text{logit}(\pi_{i\ell m}) & = \mu + \alpha_{i} + \theta\mathbb{I}(x_{\ell 1} = 1) + \gamma x_{\ell 2} + \\ & (\alpha\theta)_i \mathbb{I}(x_{\ell 1} = 1) + (\alpha\gamma)_i x_{\ell 2} + (\theta\gamma)\mathbb{I}(x_{\ell 1} = 1)x_{\ell 2} + \\ & (\alpha\theta\gamma)_i\mathbb{I}(x_{\ell 1} = 1) x_{\ell 2} + \delta_{\ell} + \epsilon_{m} \label{eq:glmm} \\
\delta_{\ell} & \sim N(0, \sigma^2_{\delta}) \\
\epsilon_{m} & \sim N(0, \sigma^2_{\epsilon})
\end{split}
\end{align}
The results of fitting this model using \texttt{glmer} from the \texttt{lme4} package are summarized in Table~\ref{tab:glmmests} \citep{lme4}. The baseline condition for this model against which all other experimental conditions are compared is the first condition in Table~\ref{tab:experdetail}, for parameter $\beta_1$ and lineup type -1. The expected value of the link function for this scenario with $\beta_1 = 0$ is $\mu$, and the expected probability that a new observer will pick out the data plot in a new lineup generated from this scenario is $\frac{\exp(\mu)}{1+\exp{\mu}}$. Similarly, for a new lineup generated from the tenth condition in in Table~\ref{tab:experdetail}, with $\beta_6 = 1$, and lineup type 1, the expected value of the link function is $\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4$. The corresponding expected probability that a new observer viewing a new lineup from this scenario is: 
\begin{equation}
\frac{\exp\{\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4\}}{1 + \exp\{\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4\}}
\end{equation}

@@

<<predictglmm, fig.height=6, fig.cap='Predictions from our generalized linear mixed effects model given in Equation~\ref{eq:glmm}. The lines show the expected probability of detecting the alternative data in a lineup of size 6 for new observers of new lineups is plotted on the $y$-axis, and the size of the parameter of interest is on the $x$-axis. The proportions detected by our Turk participants for each lineup group are shown by the points, with the probability of picking out the data plot at random shown by a horizontal line at 1/6. The lineup marked as ``outlier" was removed from modeling. The panel for the reciprocity parameter, $\\beta_2$ is also presented in Figure~\\ref{fig:recipzoom} in more detail.'>>=
newdata <- read_csv("data/newdata_pred_glmm.csv")
newdata$param <- newdata$test_param
newdata$param <- as.factor(newdata$test_param)
levels(newdata$param) <- c("beta[1]", "beta[3]", "beta[4]", "beta[2]", "beta[6]", "beta[5]")
newdata$param <-as.character(newdata$param)
plotdata <- turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
                       type2 = ifelse(type == "one", type2, -1)) %>% 
  group_by(pic_id, size, test_param, sign, type2,alt_model) %>% summarize(
    datapick = mean(datapick)
)
plotdata$param <- plotdata$test_param
plotdata$param <- as.factor(plotdata$param)
levels(plotdata$param) <- c("beta[1]", "beta[3]", "beta[4]", "beta[2]", "beta[6]", "beta[5]")
plotdata$param <- as.character(plotdata$param)
labdat <- data_frame(test_param = "jttp", param = "beta[3]", type2 = -1, x = -5, y = .9, label = "jttp_neg_hard_2")

@@

<<recipzoom, fig.height=2.5, fig.cap='The top middle panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail. The square root of the estimate is shown on the $x$-axis. For this parameter, as its value approaches zero, the probability of identifying the alternate data model decreases, then increases, which is noticeably different from the pattern exhibited by the others. Again, a horizontal line is drawn at 1/6, the chance of selecting the data plot at random.'>>=
ggplot() + 
  geom_line(data = newdata %>% filter(test_param == "recip"), aes(x = sqrt(size), y = predictfinal, color = as.factor(type2))) + 
  geom_point(data = plotdata %>% filter(test_param == "recip"), aes(x = sqrt(size), y = datapick, color = as.factor(type2))) + 
  scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
  geom_hline(yintercept = 1/6, linetype='dashed') + 
  facet_wrap(~type2, scales = 'free_x') + 
  labs(x = "Square root of parameter value (recip)", y = "Expected probability of detection\nin a lineup of size 6") + 
  ThemeNoNet + 
  theme(legend.position = 'bottom')
@

@@

<<beta41, fig.height = 4, fig.cap='In our experiment, 52.8\\% of viewers of this plot selected the plot from the alternative model, M4. The ``reverse" of this lineup is given in Figure~\\ref{fig:beta4neg1}, where 41.4\\% of viewers selected the plot from the alternative model, M1. Here, the alternative plot is $\\sqrt{25} - 3$.'>>=
dat <- read_csv("data/jtts_pos_easy_3.csv")
# "answer" is 2
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
@

<<beta4neg1,fig.height=4, fig.cap='In our experiment, 41.4\\% of viewers of this plot selected the plot from the alternative model, M1. The ``reverse" of this lineup is given in Figure~\\ref{fig:beta41}, where 52.8\\% of viewers selected the plot from the alternative model, M1. Here, the alternative plot is $\\sqrt{25} - 1$.' >>=
dat <- read_csv("data/jtts_neg_med_2.csv")
# 'answer' is 4
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
@

@@

We have found the visual power of some effects in the object function of a SAOM for this particular senate data example,  and we have shown that, for the same effects, there is a lot of variability in results from significance and goodness of fit tests. Because the visual tests we performed show a great deal of variability, we can see that the decisions with respect to the significance of a parameter or the goodness of fit of a model to data are not as cut-and-dried as the traditional methods would have us believe.

These results do not come without limitations. In visual inference, the null plots are supposed to play the role of good representatives of the null model. Here, the number of null plots is reduced to five, which increases the variability seen from a single lineup dramatically, and can unfortunately lead to different conclusions for the same lineup scenario. Furthermore, these results do not generalize yet. The lineups shown are made for only one test case, and it is not clear whether the power results transfer, nor is it clear to what degree if they do transfer, to other situations with different number of actors and different edge density.
